Automatic conversion of colloquial Finnish to standard Finnish

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
some improvements
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58605 72bbbca6-d526-0410-a7d9-f06f51895060

\begin{table*}
  \centering
  \begin{tabular}{|l|r|l|l|}
    \hline
    \textbf{Input:} & \multicolumn{3}{l|}{M\"{a} oon Tomminkaa `I am with Tommi'.} \\
    \hline
    \multirow{2}{*}{Step 1} & & M\"{a} oon Tomminkaa & apply rule 1: \\
                            & & Min\"{a} oon Tomminkaa & m\"{a} $\rightarrow$ min\"{a} \\
    \hline
    \multirow{2}{*}{Step 2} & & M\"{a} oon Tomminkaa & apply rule 2: \\
                            &   & Min\"{a} oon Tomminkaa & oon $\rightarrow$ olen \\
                            &   & M\"{a} olen Tomminkaa & \\
                            &   & Min\"{a} olen Tomminkaa & \\
    \hline
    \multirow{2}{*}{Step 3} &   & M\"{a} oon Tomminkaa   & apply rule 3:\\
                            &   & Min\"{a} oon Tomminkaa  & (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
                            &   & M\"{a} olen Tomminkaa  &\\
                            &   & Min\"{a} olen Tomminkaa  &\\
                            &   & M\"{a} oon Tommin kanssa  &\\
                            &   & Min\"{a} oon Tommin kanssa  &\\
                            &   & M\"{a} olen Tommin kanssa  &\\
                            &   & Min\"{a} olen Tommin kanssa  &\\
    \hline
    \multirow{2}{*}{Step 4} & -4.5811 &  Min\"{a} olen Tommin kanssa & rank candidates \\
                            & -7.8174 &  Min\"{a} oon Tommin kanssa & \\
                            & -8.0941 &  M\"{a} olen Tommin kanssa & \\
                            & -8.8651&  M\"{a} oon Tommin kanssa & \\
                            & -9.2045 &  Min\"{a} olen Tomminkaa & \\
                            & -12.4408 &  Min\"{a} oon Tomminkaa & \\
                            & -12.7176 &  M\"{a} olen Tomminkaa & \\
                            & -13.4885 &  M\"{a} oon Tomminkaa & \\

@@

% unsupervised does as well as CBMT, and and half the way towards pbmt

@@

% the rules written from the development portion of the corpus outperform the rules from the grammar

@@

%colloq: ooh what onks there snow
%pbmt: ooh what if there's snow
%ref: ooh what, is there snow?

>>>>

\begin{table*}
  \centering
  \begin{tabular}{|l|l|r|l|}
    \hline
    \textbf{Input:} & \multicolumn{3}{l|}{M\"{a} oon Tomminkaa `I am with Tommi'.} \\
    \hline
    \multirow{2}{*}{Step 1}  & M\"{a} oon Tomminkaa & & apply rule 1: \\
                             & Min\"{a} oon Tomminkaa & & m\"{a} $\rightarrow$ min\"{a} \\
    \hline
    \multirow{2}{*}{Step 2}  & M\"{a} oon Tomminkaa & & apply rule 2: \\
                             & Min\"{a} oon Tomminkaa & & oon $\rightarrow$ olen \\
                             & M\"{a} olen Tomminkaa & & \\
                             & Min\"{a} olen Tomminkaa & & \\
    \hline
    \multirow{2}{*}{Step 3}  & M\"{a} oon Tomminkaa   & & apply rule 3:\\
                             & Min\"{a} oon Tomminkaa &  & (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
                             & M\"{a} olen Tomminkaa  & &\\
                             & Min\"{a} olen Tomminkaa  & &\\
                             & M\"{a} oon Tommin kanssa  & &\\
                             & Min\"{a} oon Tommin kanssa  & &\\
                             & M\"{a} olen Tommin kanssa  & &\\
                             & Min\"{a} olen Tommin kanssa  & &\\
    \hline
    \multirow{2}{*}{Step 4} &  Min\"{a} olen Tommin kanssa & -4.5811 & rank candidates \\
                            &  Min\"{a} oon Tommin kanssa & -7.8174 & \\
                            &  M\"{a} olen Tommin kanssa & -8.0941 & \\
                            &  M\"{a} oon Tommin kanssa & -8.8651 & \\
                            &  Min\"{a} olen Tomminkaa & -9.2045 & \\
                            &  Min\"{a} oon Tomminkaa & -12.4408 & \\
                            &  M\"{a} olen Tomminkaa & -12.7176 & \\
                            &  M\"{a} oon Tomminkaa & -13.4885 & \\

@@

The baseline was made by calculating the metric scores between the standardised `reference' 
and the colloquial input. The results show that all conversion methods outperform the baseline.

Our unsupervised method performs similarly to the character-based machine translation systems. Both 
the character-based systems and the unsupervised system achieve around half of the performance of the 
supervised phrase-based system.

Out of the unsupervised systems, the set of rules which was created by examining the development corpus
outperforms both the set of rules from the grammar and the combined rules.

@@

@@

%! ooh mitä onks siellä lunta
%ref: ooh what, is there snow?
%colloq: ooh what onks there snow
%book: what! Is there snow?
%pbmt: ooh what if there's snow
%rules: ooh what if there's snow

%! mä vaa protestoin et ei siinä mitää tavurajaa oo i:den välissä
%ref: I was just protesting that there's no syllable break between the two i's
%colloq: I VAA you do not protest it makes no tavurajaa oo i: in between
%book: I just protest, that's not anything tavurajaa no i in between
%pbmt: I only protest that there is no tavurajaa in between the i-den
%rules: I VAA you do not protest it makes no tavurajaa no i in between

%! (kyse oli siis vaan jostain 3 rivistä loopin sisällä)
% ref: (it was only like 3 lines inside a loop)
% colloq: (it was, therefore, but for some three lines within the loop)
% book: (it was, therefore, only three to a row within a loop)
% pbmt: (it was, therefore, only some three-line within the loop)
% rules: (it was Siisi only some 3 lines of the loop inside)
